Conversations in time: interactive visualisation to explore structured temporal data

### 2020-08-10 tsibbletalk.Rmd
reworked on tidyverts intro

---
title: "Conversations in time: interactive visualization to explore structured temporal data"
author:
- name: Earo Wang
  affiliation: The University of Auckland
  address:
  - Department of Statistics
  - The Univerisity of Auckland
  - New Zealand
  email: earo.wang@auckland.ac.nz
- name: Dianne Cook
  affiliation: Monash University
  address:
  - Department of Econometrics and Business Statistics
  - Monash University
  - Australia
  email: dicook@monash.edu
bibliography: references.bib
abstract: >
  An abstract of less than 150 words.
preamble: |
  % Any extra LaTeX you need in the preamble
output: rticles::rjournal_article
---

```{r setup, echo = FALSE, cache = FALSE, include = FALSE}
options("knitr.graphics.auto_pdf" = TRUE)
library(knitr)
opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
  fig.path = 'figure/', fig.align = 'center', fig.show = 'hold',
  cache = TRUE, cache.path = 'cache/',
  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
)
knitr::opts_knit$set(root.dir = here::here())
```

@@

## Interactive visualization systems with linking views

@@

The \CRANpkg{tsibble} package [@wang2020tsibble] introduces a unified temporal data structure, referred to as a `tsibble`, to represent time series and longitudinal data in a tidy format [@wickham2014tidy]. That said, a `tsibble` extends the `data.frame` and \CRANpkg{tibble} class with temporally contextual metadata: `index` and `key`. The `index` declares a data column that holds time-related indices. The `key` identifies a collection of related series or panels observed over the `index`-defined period, which can contain multiple columns. Below displays the monthly Australian retail trade turnover data, available in the \CRANpkg{tsibbledata} package. The `Month` column holds year-months as `index`. The `State` together with `Industry` are the identifiers for these 152 series, highlighted as `key`. Note that the column `Series ID` could be an alternative option for setting up `key`, but `State` and `Industry` are informative. The `index` and `key` are "sticky" columns to a `tsibble`, forming critical pieces for fluent temporal data analysis later.

```{r, aus-retail}
library(tsibble)
library(tsibbledata)
print(aus_retail, n = 5)
```

The \CRANpkg{feasts} and \CRANpkg{fable} packages, part of the **tidyverts** suite, aim to make time series analysis easier. They provide analytical and forecasting tools for the tsibble data structure, generating tsibble-centered workflow. Functions, such as `features()` and `model()`, summarise a sequence of indexed values down to a single statistic or model by every observational unit. The output is a normal table, where each row corresponds to an observational unit denoted by "key". In the context of relational databases, the "key" acts like a foreign key in a reduced form of tsibble, while the index and key together operates like a primary key.

>>>>

---
title: "Conversations in time: interactive visualisation to explore structured temporal data"
author:
- name: Earo Wang
  affiliation: The University of Auckland
  address:
  - Department of Statistics
  - The Univerisity of Auckland
  - New Zealand
  email: earo.wang@auckland.ac.nz
- name: Dianne Cook
  affiliation: Monash University
  address:
  - Department of Econometrics and Business Statistics
  - Monash University
  - Australia
  email: dicook@monash.edu
bibliography: references.bib
abstract: >
  An abstract of less than 150 words.
preamble: |
  % Any extra LaTeX you need in the preamble
output: rticles::rjournal_article
---

```{r setup, echo = FALSE, cache = FALSE, include = FALSE}
options("knitr.graphics.auto_pdf" = TRUE)
library(knitr)
opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
  fig.path = 'figure/', fig.align = 'center', fig.show = 'hold',
  cache = TRUE, cache.path = 'cache/',
  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
)
opts_knit$set(root.dir = here::here())
read_chunk("scripts/demo.R")
```

```{r, load-pkgs}
```

@@

## Interactive visualisation systems with linking views

@@

The \CRANpkg{tsibble} package [@wang2020tsibble] introduces a unified temporal data structure, referred to as a `tsibble`, to represent time series and longitudinal data in a tidy format [@wickham2014tidy]. That said, a `tsibble` extends the `data.frame` and \CRANpkg{tibble} class with temporally contextual metadata: `index` and `key`. The `index` declares a data column that holds time-related indices. The `key` identifies a collection of related series or panels observed over the `index`-defined period, which can contain multiple columns. Below displays the monthly Australian retail trade turnover data, available in the \CRANpkg{tsibbledata} package. The `Month` column holds year-months as `index`. The `State` together with `Industry` are the identifiers for these 152 series, highlighted as `key`. Note that the column `Series ID` could be an alternative option for setting up `key`, but `State` and `Industry` are more readable and informative. The `index` and `key` are "sticky" columns to a `tsibble`, forming critical pieces for fluent temporal data analysis later.

```{r, print-retail}
```

In the spirit of tidy data to the \CRANpkg{tidyverse} [@Wickham2019], the **tidyverts** suite features `tsibble` as the foundational data structure, to build a fluid and fluent pipeline for time series analysis. Besides \CRANpkg{tsibble}, the \CRANpkg{feasts} and \CRANpkg{fable} packages fill the role of statistical analysis and forecasting in the **tidyverts** ecosystem. When time series analysis starts taking off, series of interests defined by `key` are likely to remain unchanged over the course of analysis. Functions, such as `features()` and `model()`, summarise a sequence of indexed values down to a single statistic or model by every observational unit. The output is a normal table, where each row corresponds to an observational unit denoted by "key". In the context of relational databases, the "key" acts like a foreign key in a reduced form of tsibble, while the index and key together operates like a primary key. (tsibble-centered workflow)

```{r highlight-retail, fig.height = 3.3, fig.cap = "ToDo"}
```

### 2020-08-10 tsibbletalk.tex
reworked on tidyverts intro

% !TeX root = RJwrapper.tex
\title{Conversations in time: interactive visualization to explore structured
temporal data}
\author{by Earo Wang, Dianne Cook}

@@

\hypertarget{interactive-visualization-systems-with-linking-views}{%
\subsection{Interactive visualization systems with linking
views}\label{interactive-visualization-systems-with-linking-views}}

@@

The \CRANpkg{tsibble} package \citep{wang2020tsibble} introduces a
unified temporal data structure, referred to as a \texttt{tsibble}, to
represent time series and longitudinal data in a tidy format
\citep{wickham2014tidy}. That said, a \texttt{tsibble} extends the
\texttt{data.frame} and \CRANpkg{tibble} class with temporally
contextual metadata: \texttt{index} and \texttt{key}. The \texttt{index}
declares a data column that holds time-related indices. The \texttt{key}
identifies a collection of related series or panels observed over the
\texttt{index}-defined period, which can contain multiple columns. Below
displays the monthly Australian retail trade turnover data, available in
the \CRANpkg{tsibbledata} package. The \texttt{Month} column holds
year-months as \texttt{index}. The \texttt{State} together with
\texttt{Industry} are the identifiers for these 152 series, highlighted
as \texttt{key}. Note that the column \texttt{Series\ ID} could be an
alternative option for setting up \texttt{key}, but \texttt{State} and
\texttt{Industry} are informative. The \texttt{index} and \texttt{key}
are ``sticky'' columns to a \texttt{tsibble}, forming critical pieces
for fluent temporal data analysis.

@@

The \CRANpkg{feasts} and \CRANpkg{fable} packages, part of the
\textbf{tidyverts} suite, aim to make time series analysis easier. They
provide analytical and forecasting tools for the tsibble data structure,
generating tsibble-centered workflow. Functions, such as
\texttt{features()} and \texttt{model()}, summarise a sequence of
indexed values down to a single statistic or model by every
observational unit. The output is a normal table, where each row
corresponds to an observational unit denoted by ``key''. In the context
of relational databases, the ``key'' acts like a foreign key in a
reduced form of tsibble, while the index and key together operates like
a primary key.

>>>>

% !TeX root = RJwrapper.tex
\title{Conversations in time: interactive visualisation to explore structured
temporal data}
\author{by Earo Wang, Dianne Cook}

@@

\hypertarget{interactive-visualisation-systems-with-linking-views}{%
\subsection{Interactive visualisation systems with linking
views}\label{interactive-visualisation-systems-with-linking-views}}

@@

The \CRANpkg{tsibble} package \citep{wang2020tsibble} introduces a
unified temporal data structure, referred to as a \texttt{tsibble}, to
represent time series and longitudinal data in a tidy format
\citep{wickham2014tidy}. That said, a \texttt{tsibble} extends the
\texttt{data.frame} and \CRANpkg{tibble} class with temporally
contextual metadata: \texttt{index} and \texttt{key}. The \texttt{index}
declares a data column that holds time-related indices. The \texttt{key}
identifies a collection of related series or panels observed over the
\texttt{index}-defined period, which can contain multiple columns. Below
displays the monthly Australian retail trade turnover data, available in
the \CRANpkg{tsibbledata} package. The \texttt{Month} column holds
year-months as \texttt{index}. The \texttt{State} together with
\texttt{Industry} are the identifiers for these 152 series, highlighted
as \texttt{key}. Note that the column \texttt{Series\ ID} could be an
alternative option for setting up \texttt{key}, but \texttt{State} and
\texttt{Industry} are more readable and informative. The \texttt{index}
and \texttt{key} are ``sticky'' columns to a \texttt{tsibble}, forming
critical pieces for fluent temporal data analysis later.

@@

In the spirit of tidy data to the \CRANpkg{tidyverse}
\citep{Wickham2019}, the \textbf{tidyverts} suite features
\texttt{tsibble} as the foundational data structure, to build a fluid
and fluent pipeline for time series analysis. Besides \CRANpkg{tsibble},
the \CRANpkg{feasts} and \CRANpkg{fable} packages fill the role of
statistical analysis and forecasting in the \textbf{tidyverts}
ecosystem. When time series analysis starts taking off, series of
interests defined by \texttt{key} are likely to remain unchanged over
the course of analysis. Functions, such as \texttt{features()} and
\texttt{model()}, summarise a sequence of indexed values down to a
single statistic or model by every observational unit. The output is a
normal table, where each row corresponds to an observational unit
denoted by ``key''. In the context of relational databases, the ``key''
acts like a foreign key in a reduced form of tsibble, while the index
and key together operates like a primary key. (tsibble-centered
workflow)

\begin{Schunk}
\begin{figure}

{\centering \includegraphics[width=\textwidth]{figure/highlight-retail-1}

}

\caption[ToDo]{ToDo}\label{fig:highlight-retail}
\end{figure}
\end{Schunk}
